
% to choose your degree
% please un-comment just one of the following
\documentclass[bsc,frontabs,twoside,singlespacing,parskip,deptreport]{infthesis}     % for BSc, BEng etc.
\usepackage[T1]{fontenc}
\usepackage{amssymb,amsmath}
\usepackage{txfonts}
\usepackage{microtype}

% For figures
\usepackage{graphicx}
\usepackage{subfigure} 

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% the hyperref package is used to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{mlp2017} with
% \usepackage[nohyperref]{mlp2017} below.
\usepackage{hyperref}
\usepackage{url}
\urlstyle{same}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}
\begin{document}

\title{Why is AI "a sea of dudes"? Using data science and NLP methods to understand gender imbalance in a scientific community.}

\author{Ramona Comanescu}

% to choose your course
% please un-comment just one of the following
\course{Computer Science}

% please un-comment just one of the following
\project{Undergraduate Dissertation} % CS&E, E&SE, AI&L

\project{4th Year Project Report}


\maketitle

\section*{Acknowledgements}
% I would like to thank my superviser, Dr. Adam Lopez for his unvaluable time and guidance during my project.
% My parents for the unexpected help with labeling names for gender and for the immense support. 
%Francisco Vargas for the insightful brainstorming, especially regarding name classification and helping me understand details of LDA.


\chapter{Introduction}
\section{Motivation}
Artificial intelligence (AI) has been growing rapidly in recent years. However, just 13\% of one of 2015's biggest AI conferences (NIPS) were women, according to \cite{seaofdudes}
Microsoft researcher Margaret Mitchell calls Artificial Intelligence "a sea of dudes" and this situation is definitely concerning. If you train a speech recognition system using British English, it will not generalize well when you test it on American English. \\
There are some famous examples of how biased or incomplete datasets create undesirable results: Tay, Microsoft chat-bot learned from users' tweets to be racist \citep{tay}, while Google mistakenly tagged black people as gorillas\citep{gorilla}. Machine learning can amplify biases already present in the data and researchers are trying to investigate this. \cite{homemaker} shows that continuous vector space representation of word embeddings trained on Google News articles learn the hidden sexism in language, resulting in interesting vector space arithmetic:
\begin{equation*}
\overrightarrow{man} - \overrightarrow{woman} \approx \overrightarrow{computer programmer} - \overrightarrow{housemaker}
\end{equation*}
What is more, bias is not only about training data, machine learning systems also suffer if the designers do not come from diverse backgrounds, specifically if they do not account for gender. We aim to investigate the gender from the perspective of those shaping the future of AI.

\section{Outline}

The main goal of this thesis is to analyse gender imbalance in the field of computational linguistics. We go beyond surface statistics and try to understand how gender affects academia, from the output of research to the career paths of individuals. We formulate different tasks and use data science and NLP methods to understand them. Topics and methodologies we cover are outlined below:

\section{Summary of contributions}
 All investigations carried are written as general Python modules and are corpus agnostic, once the corpus is standardized into a predefined csv format.
\begin{itemize}
\item{Data collection}
\item{
    \begin{itemize} 
    \item{Cleaning up AAN corpus (removing duplicates, identifying      malformed documents, standardizing escape characters)}
    \item{Scraping ArXiv for Computational Linguistics Papers}
    \end{itemize}
}
\item{Name classification
 \be}
Publishing patterns:
Topic modeling:
 Preprocessing the data into bag of words, using Gensim and Spacy tokenizers, lemmatizers, named entity recognition tools
 Constructing LDA models and comparing perplexity of models with different parameters on a validation set.
 \end{itemize}
 

\begin{itemize}
\item 
\item In chapter \ref{ch:publ} we compare publishing patterns of men and women. 
    \begin{itemize}
    \item Single authored papers and coauthorship
    \item Collaboration network
    \item 
    \end{itemize}
\end{itemize}

\chapter{Background}

\section{Previous work}

Most studies of gender balance in computer science are based on reporting statistics on enrolment. \cite{Jurafsky_Hesaid} go beyond the surface and study the difference in topics that different genders write about. They perform a "mostly-manual annotation" of the gender of each author, which we also use as a starting point for our name classification. They use Latent Dirichlet Allocation(LDA) topic models \citep{Blei_LDA} to study the difference in topics genders write about. Their study is conducted on the Association of Computational Linguistics Anthology Network (AAN) corpus \citep{aan}, using papers from 1965 to 2008. They find that "women publish more on dialogue, discourse, and sentiment, while men publish more than women in parsing, formal semantics and finite state models". They also find that the participation of females in the field has been steadily increasing.

 The resulting picture closely resembles and adds to that arrived at through qualitative analysis, showing that this form of topic modeling could be useful for sifting through datasets that had not previously been subject to any analysis. The strength of the relationships between topics illustrates the fluid and ubiquitous nature of sexism, with no single experience being unrelated to another.

Topic models for text corpora comprise a popular family of methods that have inspired many extensions to encode properties such as sparsity, interactions with covariates, and the gradual evolution of topics.

\section{Why gender balance matters}

\section{Ethical concerns}

\chapter{Methodology}

\section{Corpus}
\subsection{ACL Anthology Network Corpus}
We use The ACL anthology network corpus \cite{aan} with the lastest release being 2014. 

The  ACL 2012 Contributed Task \cite{contributed} outlines some of the problems that make maintaining an up to date and error-free ACL collection difficult. We make our own attempt at cleaning and preprocessing the data, before starting doing any analysis.
We remove all duplicate ids(papers published in multiple publication) and we are left with 22944 papers. 

The source dataset describes both automatic conversion from pdf to text and OCR techniques. However, we discover files for which conversion has failed, spitting out random character.
In order to detect malfunctioning conversion, we do automatic language detection, using \textbf{langid.py.} \cite{langid}
Pdf to text conversion only works for born digital pdf papers. Once we identify the files for which pdftotext failed, we used OCR with Tesseract. In the process, we also discover some papers where in french and we need to discard them. We eliminate 515 corrupted papers.
\subsection{ArXiv Corpus}

\section{Name classification}
\subsection{Methods}
\subsection{Evaluation}

\chapter{Publishing patterns}
\label{ch:publ}
\section{Summary statistics}
\section{Single authored papers and coauthorship}
\section{Collaboration network}

\chapter{Topic modelling}
\label{ch:topics}
\section{Latent Dirichlet Allocation}
% what it is
\section{Author-topic models}
\section{Experiments}
\section{Results}

\chapter{Stylometry}
\label{ch:style}
\section{Features}

\chapter{Conclusion}
As there is limited quantitative analysis of women and
men's scholarly output computer science(most analysis just report statistics on the number of graduates), this project
fills a gap in our knowledge about sex and publication productivity. We also analysed topics in research papers based on gender.

\tableofcontents

%\pagenumbering{arabic}

\bibliographystyle{apa}
\bibliography{mybibfile}


\end{document}

